Breadth First Searches Over Evolving Graphs

### 2016-01-26 paper.tex
update acknowledgements
fix typo
Fix language and M matrix definition

We can use analyze the spread of information between the players using graph
theory. In this process, the time ordering of events matters, and hence its
graph representation $G(t) = (V(t), E(t))$ must be time dependent.
Such a graph is called an ``evolving graph''~\cite{fkp89,bkmu12},
``evolving network'' or %XXX WHO?
``time-dependent graph''. %XXX WHO?

@@

\begin{defn}
An \textbf{active node} is a temporal node $(v,t)$ if there exists at least one
edge $e\in E^{[t]}$ that connects $v\in V^{[t]}$ to another node $w\in V^{[t]}$,
$w\ne v$.

@@

This definition of a temporal path differs from that of the dynamic walk
in \cite{gphe11,grihig13} in its explicit requirement of containing only active
nodes. The definition implies that if either or both end points of a temporal
path are inactive, then the entire temporal path must be the empty sequence
$\langle \rangle$. Keeping track explicitly of the time labels of each temporal
node allows greater generality to cases where the node sets change over time.
Furthermore, we shall show later in \ref{sec:temp-paths-adjac}
that the explicit bookkeeping of the time labels is essential for correctly
generalizing the BFS to evolving graphs.

@@

In Figure~\ref{fig:eg_shortest_path},
the forward neighbors of $(1, t_1)$ are $(2, t_1)$ and $(1, t_2)$ and
the forward neighbors of $(2, t_1)$ is $(2, t_3)$.
The 2-forward neighbors of $(1, t_1)$ are
$(2, t_1), (1, t_2), (2, t_2)$ and $(3, t_2)$.
By construction, time stamp of every forward neighbor of an active node $(v, t)$
must be no earlier than $t$.

@@

\IEEEproof
Define the set of temporal nodes
$\tilde V_L^{[t]} = \{(v_1)_t = (v_1, t) | (v_1, v_2) \in E^{[t]}\}$, which
consists of the active nodes at time $t$ which participate on the left side of an edge.
Similarly, $\tilde V_R^{[t]} = \{(v_2)_t = (v_2, t) | (v_1, v_2) \in E^{[t]}\}$
are the corresponding active nodes on the right side of an edge. Then
$\tilde V^{[t]} = \tilde V_L^{[t]} \cup \tilde V_R^{[t]}$ is the set of active
nodes at time $t$, and $V = \bigcup_t \tilde V^{[t]}$ is the set of all
active nodes in $G_n$.

@@

Sums $S^{[t]}$ of the form \eqref{eq:sum} produce an incorrect count of temporal
paths because they do not capture temporal paths with subpaths of the form
$\langle (v, s), (v, t) \rangle$, $s < t$. The temporal path \eqref{eq:tpathex}
is counted only by the matrix product $M^{[t_1]} A^{[t_2]}$, where
\[
    M^{[t_1]} = \begin{pmatrix}1 & 0 & 0 \\ 0 & 1 & 0 \\ 0 & 0 & 0\end{pmatrix}.
\]
$M^{[t_1]}$ describes the forward time propagation of active nodes. It counts
temporal paths with subpaths $\langle (i, t_1), (i, t_2) \rangle$
for all active nodes $(i, t_1)$ at time $t_1$, but not paths containing the
subsequence $\langle (3, t_1), (3, t_2) \rangle$, which cannot exist in temporal
paths since $(3, t_1)$ is an inactive node.

One might attempt to amend the sums $S^{[t]}$ in \eqref{eq:sum} by redefining
the adjacency matrices to include ones along the diagonal. However, the resulting
sum is still incorrect, as it counts paths with subsequences $\langle (3, t_1), (3, t_2) \rangle$
and are hence not temporal paths.

@@

We thank N. J. Higham and V. \u Sego (Manchester) for helpful suggestions.
W.\ Z.\ thanks the School of Mathematics at U. Manchester for travel funding and
A. Edelman for arranging for a fruitful visit to MIT.

%XXX ANY OTHER FUNDING

>>>>

We can analyze the spread of information between the players using graph
theory. In this process, the time ordering of events matters, and hence its
graph representation $G(t) = (V(t), E(t))$ must be time dependent.
Such a graph is called an ``evolving graph''~\cite{fkp89,bkmu12},
``evolving network'' or %XXX WHO?
``time-dependent graph''. %XXX WHO?

@@

\begin{defn}
A temporal node $(v,t)$ is an \textbf{active node} is there exists at least one
edge $e\in E^{[t]}$ that connects $v\in V^{[t]}$ to another node $w\in V^{[t]}$,
$w\ne v$.

@@

This definition of a temporal path differs from that of the dynamic walk
in \cite{gphe11,grihig13} in its explicit requirement of containing only active
nodes. The definition implies that if either or both end points of a temporal
path are inactive, then the entire temporal path must be the empty sequence
$\langle \rangle$. Keeping track explicitly of the time labels of each temporal
node allows greater generality to cases where the node sets change over time.
Furthermore, we shall show later in Sec.~\ref{sec:temp-paths-adjac}
that the explicit bookkeeping of the time labels is essential for correctly
generalizing the BFS to evolving graphs.

@@

In Figure~\ref{fig:eg_shortest_path},
the forward neighbors of $(1, t_1)$ are $(2, t_1)$ and $(1, t_2)$ and
the only forward neighbor of $(2, t_1)$ is $(2, t_3)$.
The 2-forward neighbors of $(1, t_1)$ are
$(2, t_1), (1, t_2), (2, t_2)$ and $(3, t_2)$.
By construction, time stamp of every forward neighbor of an active node $(v, t)$
must be no earlier than $t$.

@@

\IEEEproof
Define the set of temporal nodes
$\tilde V_L^{[t]} = \{(v_1, t) | (v_1, v_2) \in E^{[t]}\}$, which
consists of the active nodes at time $t$ which participate on the left side of an edge.
Similarly, $\tilde V_R^{[t]} = \{ (v_2, t) | (v_1, v_2) \in E^{[t]}\}$
are the corresponding active nodes on the right side of an edge. Then
$\tilde V^{[t]} = \tilde V_L^{[t]} \cup \tilde V_R^{[t]}$ is the set of active
nodes at time $t$, and $V = \bigcup_t \tilde V^{[t]}$ is the set of all
active nodes in $G_n$.

@@

Sums $S^{[t]}$ of the form \eqref{eq:sum} produce an incorrect count of temporal
paths because they do not capture temporal paths with subpaths of the form
$\langle (v, s), (v, t) \rangle$, $s < t$.
One might attempt to amend the sums $S^{[t]}$ in \eqref{eq:sum} by redefining
the adjacency matrices to include ones along the diagonal, hence allowing
paths containing the sequence $\langle (i, t_1), (i, t_2) \rangle$. However, the resulting
sum is still incorrect, as it counts paths with subsequences $\langle (3, t_1), (3, t_2) \rangle$
and are hence not temporal paths. Instead, the temporal path \eqref{eq:tpathex}
is counted only by the matrix product $M^{[t_1, t_2]} A^{[t_2]}$, where
\[
    M^{[t_1, t_2]} = \begin{pmatrix}1 & 0 & 0 \\ 0 & 0 & 0 \\ 0 & 0 & 0\end{pmatrix}.
\]
$M^{[t_1, t_2]}$ describes the forward time propagation of temporals nodes that
are active at both times $t_1$ and $t_2$, i.e.\ it counts
temporal paths that contain subsequences $\langle (i, t_1), (i, t_2) \rangle$,
and both $(i, t_1)$ and $(i, t_2)$ are active nodes.

@@

We thank N. J. Higham and V. \u Sego (Manchester) for helpful suggestions.
W.\ Z.\ thanks the School of Mathematics at U. Manchester for research \& travel funding and
A. Edelman for arranging for a fruitful visit to MIT.
